---
title: "Lab2_Bologna"
author: "Justin Bologna"
date: "April 18, 2020"
output: html_document
---
##Lab 2 - Descriptive Statistics and Basic Mapping

Today we are going to be examining a dataset of selected socioeconomic indicators in Chicago form the period 
2008 - 2012. The file was downloaded as a non-spatial CSV from the Chicago Data Portal website.  

First, let's set up our R environment and load the packages we need.

```{r, message=FALSE}
setwd("~/UChicago 19-20/Spring/GIS III/wd/lab2")
library(sf)
library(dplyr)
library(spData)
```


Since our dataset of interest is non-spatial, we must joing it to a spatial shapefile of Chicago community areas. 
Let's read these files into our R environment and join them.

```{r}
indicators <- read.csv("selected_indicators.csv", stringsAsFactors = FALSE)
ca <- st_read("ca.shp")

indicators[76,2] <- "OHARE"
indicators[,2] <- toupper(indicators[,2])
indicators <- rename(indicators, community = COMMUNITY.AREA.NAME)

indicators_spatial <- left_join(ca, indicators, by="community")
```

We had to tweak the names of some columns and fields to get them to match before we could join, but now we have a
spatial version of the data. Let's examine the summary statistics for a few select variables.

```{r}
summary(indicators_spatial[,c(14,16,17)])
```

Next, let's focus on one specific indicator: the hardship index. We can generate a histogram with the following code:

```{r}
Hardship_Index <- indicators_spatial$HARDSHIP.INDEX
hist(Hardship_Index)
```

Since we went through the trouble of making our dataset spatial, let's map the Hardship Index too.

```{r}
plot(indicators_spatial[,17])
```

##Works Cited

Chicago Data Portal. (2014). *Census Data - Selected socioeconomic indicators in Chicago, 2008 - 2012* [CSV]. https://data.cityofchicago.org/Health-Human-Services/Census-Data-Selected-socioeconomic-indicators-in-C/kn9c-c2s2. 

Chicago Data Portal. (2018). *Boundaries - Community Areas (current)* [Shapefile]. https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas-current-/cauq-8yn6. 