---
title: "Lab3_Bologna"
author: "Justin Bologna"
date: "April 23, 2020"
output: html_document
---
##Lab 3: Mapping Divvy Bicycle Stations from an API

Today we'll be pulling a dataset straight from the web to produce a map of Divvy bicycle stations around Chicago. First, let's load our packages.

```{r, message=FALSE}
library(ggmap)
library(jsonlite)
library(dplyr)
library(rgdal)
library(tmap)
```

Now, let's read the JSON from the web, convert it to a dataframe, and clean the data a bit.

```{r}
divvy1<- fromJSON(txt="https://data.cityofchicago.org/resource/bk89-9dk7.json")
divvy2 <- divvy1 %>% select(1,3,5,7,12,13)
divvy2[,2] <- as.numeric(divvy2[,2])
divvy2[,3] <- as.numeric(divvy2[,3])
divvy2[,4] <- as.numeric(divvy2[,4])
divvy2[,5] <- as.numeric(divvy2[,5])
divvy2[,6] <- as.numeric(divvy2[,6])
head(divvy2)
```

Great. Next we need to convert the non-spatial data frame into a spatial points data frame using the latitude and longitude columns of the table. A quick plot shows that we've done it successfully. 

```{r}
divvy_spdf <- SpatialPointsDataFrame(coords = cbind(divvy2$longitude, divvy2$latitude), data = divvy2, proj4string = CRS("+init=epsg:4326"))
plot(divvy_spdf)
```

Now let's create a more polished, interactive map to show the locations of Divvy stations and how many active docks there are at each station. We will do this using tmap and leaflet, since 

```{r}
m <- tm_shape(divvy_spdf)+
  tm_dots(col='total_docks', title='Total Docks', size=0.05)+
  tm_basemap(leaflet::providers$Esri.WorldGrayCanvas, alpha=1)+
  tm_layout(title="Divvy Bicycle Stations in Chicago")
  
tmap_leaflet(m)
```